---
title: "Example for the Cultivar Matching Process"
output:
  minidown::mini_document:
    code_folding:
      source: hide
    results_folding: hide #show, hide or none
    framework: water
    theme: light
    toc: true
    toc_float: true
    toc_highlight: true
    
---

```{r setup, eval = FALSE, include = FALSE}
rmarkdown::render(
  here::here("tests", 
    "example_cultivar_matching", 
    "raw",
     "example_cultivar_matching.Rmd"), 
  output_dir = here::here("tests",
                          "example_cultivar_matching",
                          "raw",
                          "output"))
```
```{r, include = FALSE}
library(here)
knitroutputfolder = here("tests", "example_cultivar_matching", "raw", "output")
source(here("R", "dependencies.R"))
auxiliary_files <- here("tests", 
                        "example_cultivar_matching",
                        "raw",
                        "auxiliary_files")
```

# Read in raw variety names
```{r, results = 'asis', results.folding = 'show'}
var <- 
  read_csv(here("tests", 
                "test_example_files", 
                "example_cultivar_matching.csv"), 
            col_types = cols())

kable(var)
```
# Cultivar matching 
## Create unique internal id for names
Note!: Once the fuzzymatching stage is started, it is important that the 
var_id doesn't change other it will cause problems with matching up matches
with the correct names
```{r, results = 'asis', results.folding = 'show'}
names_raw <- 
  create_intid(var, 
               variety, 
               sep_aliases = "\\(", 
               crop_type, 
               nursery) %>% mutate(var_id = as.character(var_id))#, 
               #alias_col = alias)

kable(names_raw)
```
## Do exact match
```{r, results = 'asis', results.folding = 'show'}
results_exactmatch <- 
  do_exactmatch(data_intid = names_raw,
                select_before = "2021-11-22",
                match_type = "raw")




results_rename <- 
  do_exactmatch(data_intid = results_exactmatch$nomatch, 
                                match_type = "raw", 
                                rename_df = TRUE)

kable(results_rename[["match"]])
kable(results_rename[["nomatch"]])
```
## Do fuzzymatch
The fuzzy matching procedure returns the best fuzzymatches from
stringdist_join using two methods ("lv" and "jw").  
The results are filtered for these conditions:
- crop_type must match (e.g. wheat vs. barley)
- numbers at the end of a variety name must be an exact match

### Output file to manually match
Fill the is_truematch column with TRUE if fuzzymatch is a true match, otherwise fill with FALSE
```{r, results = 'asis', results.folding = 'show'}
output_fuzzymatch <- 
  find_fuzzymatch(results_rename$nomatch ,
                intid_col = "intid",
                select_before = "2021-05-26",
                knitroutputfolder = knitroutputfolder)


kable(output_fuzzymatch)
```
### Process manual matches
Read in manual check and output names with no match
Return aliases (same var_id) that had matches in the previous steps.
These aliases were not in the original database and should be added.
If no match or aliases, then generate new cultivar entries for the database.
Write out the new names to be standardized/ checked by collaborators
```{r, results = 'asis', results.folding = 'show'}
fuzzy_status <- 
  read_csv(
    here(here(auxiliary_files), "fuzzymatch_status.csv")) %>%
  mutate(var_id = as.character(var_id))


results_fuzzymatch <- 
  process_fuzzymatch(output_fuzzymatch_df = output_fuzzymatch, 
                     aux_fuzzy_status = fuzzy_status)

# Go through and standardize the names in standardize_new_names_aux.csv
# If any names are found to already be in the database, then manually
# add them to the controlled_vocab/cv_rename.csv file
# This occurs when the name is not found by fuzzymatching
results_new_names <- 
  process_std_new_names(
    output_nomatch_df = results_fuzzymatch[["nomatch"]], 
    auxiliary_files = auxiliary_files,  
    filename = "standardize_new_names_aux.csv")

```

# Collect all the matches from each step
Once the names from the process_std_new_names() step have been added to
the cv_rename file, then they should match in the cv_rename step
```{r, results.folding = NULL, results = 'asis'}
names_matches <- collect_final_matches(list(exact = results_exactmatch,
                                            rename = results_rename,
                                 fuzzy = results_fuzzymatch), names_raw)

kable(names_matches)

```
# Bind matches with the data
```{r, eval = FALSE}
# pseudocode
# data_cv_names <- left_join(data, names_matches, by = "variety)
```