---
title: "Example of the cultivar matching process"
author: "Jacqueline Tay"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output: 
rmarkdown::html_vignette:
  fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to codebook structure}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE}
library(factcuratoR)

codebook_folder <- here::here("tests/testthat/test_controlled_vocab")
example_folder <- here::here("inst/extdata/example_cultivar_matching")
knitroutputfolder = here::here(example_folder, "output")
knitroutputfolder_blends = here::here(example_folder, "output_blends")
auxiliary_files <- here::here(example_folder, "auxiliary_files")
```

# Read in raw variety names
```{r, results = 'asis', results.folding = 'show'}
var <- read_csv(here::here(example_folder, "example_cultivar_matching.csv"), 
                col_types = cols())

knitr::kable(var)
```
# Cultivar matching 
## Create unique internal id for names
Extraneous information in the variety name will make the matching process more
difficult - the names won't match by exact matching and it is less certain 
that the fuzzymatching functions will be able to detect a fuzzymatch.

Some types of information that may be stored in the variety name column
that should be removed includes:
parental cross, crop type, agronomic treatments ("fill", "sprayed", etc).

Remove extraneous information stored in the raw variety column by duplicating the 
name column and removing the extraneous information (which can usually be
accomplished by regex).  The column is duplicated because the raw variety name 
should be retained to match database matches back with the data.

Additionally, if there are blends, they should be separated from the other variety
names and matched separately.  Many of the variety matching functions have the
argument `is_blends`, which can be set to TRUE to match the blends (See the section 
below on matching blends).  The functions will help standardize the names according 
to the formatting for blends: name1;name2

<span class="important"> Note! </span> Once the fuzzy matching stage is started, it is important that the `var_id` doesn't change other it will cause problems with matching up matches with the correct names.

```{r, results = 'asis'}
names_raw <- 
  create_intid(var, 
               variety, 
               sep_aliases = "\\(", 
               crop_type, 
               nursery) %>% mutate(var_id = as.character(var_id))#, 
               #alias_col = alias)

knitr::kable(names_raw)
```
## Do exact match
The first step is to exact match to the cultivar names in the codebooks

The second step is to exact match to the names in the cv_rename.csv.
cv_rename.csv contains misspellings.
```{r, results = 'asis'}
results_exactmatch <- 
  do_exactmatch(db_folder = codebook_folder,
                data_intid = names_raw,
                select_before = "2021-11-22",
                match_type = "raw")

results_rename <- 
  do_exactmatch(db_folder = codebook_folder, 
                data_intid = results_exactmatch$nomatch, 
                match_type = "raw", 
                rename_df = TRUE)

knitr::kable(results_rename[["match"]])
knitr::kable(results_rename[["nomatch"]])
```
## Do fuzzymatch

The fuzzy matching procedure, a wrapper for `fuzzyjoin::stringdist_join()` returns the best fuzzy matches using multiple distance methods (e.g. "lv", "jw", "lcs"). 
See `??stringdist` for more information about these distance metrics.

The results are filtered for these conditions:

* The variable in `crop_type` must match (e.g. "wheat" would not match "barley")

* Numbers located at the end of a variety name or alias must be an exact match if there is more than one number present. If there is only number at the end of a variety name, the trailing digit does not need to match the rest of the name. For example "myvar1" would still match with "myvar", but "myvar123" would not match "myvar". 

The results of `find_fuzzymatch(`) are written to the file *fuzzymatch.csv* in the
directory provided by the user in the `knitroutputfolder` argument. 

```{r, results = 'asis', results.folding = 'show'}
output_fuzzymatch <- 
  find_fuzzymatch(results_rename$nomatch,
                intid_col = "intid",
                select_before = "2021-05-26",
                knitroutputfolder = knitroutputfolder,
                db_folder = codebook_folder)

knitr::kable(output_fuzzymatch)
```

## Fill out the fuzzymatch.csv

Move *fuzzymatch.csv* to the helper_files/ directory and fill out the `is_truematch` column.
(You can append the file with the ending "_aux" to clearly differentiate it between
the .csv that will be written to the output/ directory for each run of the .Rmd.)
Fill the `is_truematch` column with "TRUE" if fuzzymatch is a true match, 
otherwise fill with "FALSE".

Sometimes the match may be clear, however, many matches will need to be manually checked. Some steps for that: 

1. First, start off with a Google search to determine if there are clear usages 
of both names (name in the raw data and name in the database).  This would suggest
that the names refer to different varieties and they are not true matches.

1. If a Google search doesn't help clarify whether the name is a true match (quite likely, unfortunately), check the possible matches with our collaborators.

1. If there is no clear answer (or any answer), the curator should leave `is_truematch` blank as the subsequent steps will remind the curator to follow up on these possible matches.

## Process *fuzzymatch_aux.csv* manual matches

1. Read in the *fuzzymatch_aux.csv* that has been manually filled in
1. Process the file using `process_fuzzymatch()`, which will write out 3 files to the working directory:

  * *fuzzy_check.csv*, which contains variety names that still need to be manually 
  checked for whether the fuzzy match is a true match.  Update is_truematch in the 
  *helper_files/fuzzymatch_aux.csv* file.
  
  There may be newly detected fuzzy matches for successive runs of the .Rmd.  
  This may occur if there are new names added to the codebook or if the 
  fuzzymatching function is updated.
  
  Newly detected fuzzy matches need to be added to the *helper_files/fuzzymatch_aux.csv*
  file.  The new matches have only "new_fuzzymatch" in the source column.
  ("new_fuzzymatch;csv" in the source column means that this was a match for this
  run of the .Rmd ("new_fuzzymatch") and that the name is also already present in the original fuzzymatch.csv ("csv").)
 
  * *add_fuzzy_to_cv_rename.csv*, which contains variety names for which 
  `is_truematch` is "TRUE."  Manually add the names to *cv_rename.csv*.
  
  * *standardize_new_names.csv*, which contains variety names that have no match 
  and need to be standardized before being added to the database.  See next
  section for full details.  

```{r, results = 'asis', results.folding = 'show'}
fuzzy_status <- 
  read_csv(here::here(auxiliary_files, "fuzzymatch_status.csv")) %>%
  mutate(var_id = as.character(var_id))

results_fuzzymatch <- 
  process_fuzzymatch(output_fuzzymatch_df = output_fuzzymatch, 
                     aux_fuzzy_status = fuzzy_status,
                     knitroutputfolder = knitroutputfolder)
```

## Standardize names without a match

1. Move *standardize_new_names.csv* to the helper_files/ directory 
(You can append the file with the ending "_aux" to clearly differentiate it between
the .csv that will be written to the output/ directory for each run of the .Rmd.)

1. Go through and standardize the names in *standardize_new_names_aux.csv* 
by editing the names in the column `new_std_name`. The function 
`process_fuzzymatch()` will make an initial guess by populating `new_std_name` 
with the the raw variety name. 

<span class="important">Note on duplicate names: </span> It is especially 
important if there are different spellings of the same variety, the names 
are standardized in this file. For example, if "WA0061" and "WA61" are the 
same variety (spoiler: they are), the `new_std_name` should be the same for both entries.  
If the names are not standardized, processing the names later will create two
different entries to add to the database controlled vocabularies.

<span class="important">Note on names missed by fuzzy matching: </span> 
If any of the names in *standardize_new_names_aux.csv* match a variety in the database,
then manually add the names to the *controlled_vocab/cv_rename.csv* file. 
This occurs when the name is not found by fuzzy matching.  The fuzzy matching functions have been written
to return plausible matches while trying to minimize the number of total returns
so to not overwhelm the user (you).  This trade off means that some potential matches may be missed.

1. Read in and process the file using `process_std_new_names()`, which will write 
out files to the /output folder: 
* *add_to_new_std_name.csv.*, which contains names that have newly passed from the
fuzzy matching stage to the name processing stage.  If there are 
entries in this file, add the names to *helper_files/standardize_new_names_aux.csv*
and standardize the `std_new_name` column as described above.
* *add_to_cv_rename.csv*, which are names that have been updated in 
*standardize_new_names_aux.csv* and won't match by exact_matching (i.e. if the 
`std_new_name` was changed from the default value.  If there are 
entries in this file, add the names to *controlled_vocab/cv_rename.csv*
* *add_to_db.csv*, which contains new varieties. Add these entries to the database.
May need to manually fill in extra information such as crop or alias

```{r}
results_new_names <- 
  process_std_new_names(
    output_nomatch_df = results_fuzzymatch[["nomatch"]], 
    auxiliary_files = auxiliary_files,  
    filename = "standardize_new_names_aux.csv",
    knitroutputfolder = knitroutputfolder,
    db_folder = codebook_folder)

```

## Iterate through the process

If the fuzzy_status file is updated (that is, `is_truematch` is filled in), re-run the entire process in a clean R session and `process_fuzzymatch()` and 
`process_std_new_names()` will update the output files.  

The goal is to iterate through the process outlined above until
all of the output files (*fuzzy_check.csv*, *add_to_cv_rename.csv*, 
*add_to_new_std_name.csv*, and *add_to_new_std_name.csv*) are empty.

When this has been completed, all of the variety names should be matched and
all the necessary information has been moved into the controlled vocabularies.

Hooray! Make sure you celebrate this achievement! 

# Collect all the matches from each step

Once the names from the `process_std_new_names()` step have been added to
the *cv_rename* file, then they should match in the cv_rename step. 

The function` collect_final_matches()` is used to collect all the matches and the stage 
where they were matched. 

```{r, results.folding = NULL, results = 'asis'}
names_matches <- 
  collect_final_matches(list(exact = results_exactmatch,
                             rename = results_rename,
                             fuzzy = results_fuzzymatch), 
                        names_raw,
                        knitroutputfolder = knitroutputfolder)

knitr::kable(names_matches)

```

# Bind matches with the data

```{r, eval = FALSE}
# do not run this code (it's here for example only)
data_cv_names <- left_join(data, names_matches, by = "variety")
```

# Special instructions for matching blends
Curate the blends separately from non-blend variety names.
Below is an example for curating blends.

```{r, results = 'asis'}
blends <- read_csv(here::here("tests/testthat/test_match_variety_files", 
                              "example_blends.csv"), 
                col_types = cols())

knitr::kable(blends)
```
## Create unique internal id for names
<span class="important"> Note! </span> For blends, set the argument
`is_blends = TRUE` for each of the variety matching functions.

```{r, results = 'asis'}
blends_raw <- 
  create_intid(blends, 
               variety, 
               # Although sep_alias is used to separate 
               # an alias from the variety name, it can also be used here 
               #to separate two names in a blend
               sep_aliases = "\\/", 
               crop_type,
               is_blends = TRUE) 

knitr::kable(blends_raw)
```

## Do exact matching step
```{r, results = 'asis'}
blends_exactmatch <- 
  do_exactmatch(db_folder = codebook_folder,
                data_intid = blends_raw,
                select_before = "2021-11-22",
                match_type = "raw",
                is_blends = TRUE)

knitr::kable(blends_exactmatch[["match"]])
knitr::kable(blends_exactmatch[["nomatch"]])

```
```{r, results = 'asis'}
blends_rename <- 
  do_exactmatch(db_folder = codebook_folder, 
                data_intid = blends_exactmatch$nomatch, 
                match_type = "raw", 
                rename_df = TRUE, 
                is_blends = TRUE)

knitr::kable(blends_rename[["match"]])
knitr::kable(blends_rename[["nomatch"]])

```
## Do fuzzymatching step
```{r, results = 'asis'}
blends_fuzzymatch <- 
  find_fuzzymatch(blends_rename$nomatch,
                intid_col = "intid",
                select_before = "2021-05-26",
                knitroutputfolder = knitroutputfolder_blends,
                db_folder = codebook_folder)

knitr::kable(blends_fuzzymatch)
```
## Read in and process fuzzymatches
```{r, results = 'asis'}
blends_fuzzy_status <- 
  read_csv(here::here(auxiliary_files, "fuzzymatch_blends_status.csv")) %>%
  mutate(var_id = as.character(var_id))

blends_results_fuzzymatch <- 
  process_fuzzymatch(output_fuzzymatch_df = blends_fuzzymatch, 
                     aux_fuzzy_status = blends_fuzzy_status,
                     knitroutputfolder = knitroutputfolder_blends)

```
## Collect all matches
```{r, results = 'asis'}
blends_matches <- 
  collect_final_matches(list(exact = blends_exactmatch,
                             rename = blends_rename,
                             fuzzy = blends_results_fuzzymatch), 
                        blends_raw,
                        knitroutputfolder = knitroutputfolder_blends,
                        is_blends = TRUE)

knitr::kable(blends_matches)

```
