---
title: "Introduction to the codebook structure"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: 
rmarkdown::html_vignette:
  fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to codebook structure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
The purpose of `factcuratoR` is to provide sets of functions to help 
standardize variety testing data for the FACT program.

The goal of the FACT program curation is to generate:

trial data
trial metadata
that conform to the controlled vocabulary codebooks.

## Codebook structure

First, load `factcuratoR` and point to the main codebook (which currently must 
be named codebooks_all_db.csv for the validation functions)

```{r, message = FALSE}
library(factcuratoR)

codebook_path <- here::here(
  "tests/testthat/test_controlled_vocab/codebooks_all_db.csv")
```

Most importantly, the main codebook contains:
the variables (columns) for both the trial data and the trial metadata. 
For example, the trial data should contain the columns:
trial, variety, entry, plot, etc...
and the trial metadata should contain the columns:
trial, nursery, year, location, etc...

Note! Traits are collected for each trial, so conceptually, it would make sense
for the traits (such as test_weight or height) to be contained in the main codebook
associated with trial_data. However, because the traits list is 
expected to get rather long, the traits are stored in a separate file called 
"traits.csv."  For validating the columns in the trial data, the validation 
functions pull in the traits and treats each trait as a column in the trial data.

```{r, fig.cap = "Fig. 1 Codebook structure"}
create_dm(codebook_path)
```




## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
